### 2012-11-19 doc/levelset/level_set_method_on_quadtree_JCP_rebuttal.tex

\subsection{Third order accuracy}\label{subsec_Accuracy_Reinitialization}
We also computed the convergence rates of the reinitialization algorithm
for the test problem proposed in \cite{Russo_Smereka}: Consider the level
set function initially defined as:
$$\phi^0(x,y)=\left(0.1+(x-1)^2+(y-1)^2\right)\left(\sqrt{x^2+y^2}-1\right),$$ which defines the
interface as a circle with center the origin and radius $1$. In this case, $\phi^0$ is not a signed
distance function and its gradients vary widely. Figure \ref{fig_reinitialization} illustrates the
gradual deformation of the cross-sections of $\phi^0$ as it evolves to the signed distance
function. Table \ref{tab_reinitialization} illustrates that the method is third order accurate in
the $L^1$ and $L^\infty$ norms near the interface, where we use the standard formulas for the $L^1$
and $L^\infty$ norms:
\begin{align*}
||\phi||_{\infty}=\max_{v:|\phi(v)|<1.2\Delta x}|\phi(v)-\phi_{\textrm{exact}}(v)|, \\
||\phi||_{1}=\avg_{v:|\phi(v)|<1.2\Delta x}|\phi(v)-\phi_{\textrm{exact}}(v)|, \\
\end{align*}
where $\Delta x=\Delta x_{\textrm{smallest}}$. Note that after the reinitializing the level set
function, the choice of $v:|\phi(v)|<1.2\Delta x$ ensures the selection of all the nodes adjacent
to the interface.

In the entire domain, the method is second order accurate if we keep refining all the cells. In the
practical case where only cells near the interface are refined, the accuracy in regions far away
from the interface is meaningless. In the case where the interface presents sharp corners, the
accuracy is reduced to first order in the $L^{\infty}$ norm.

%\begin{figure}[t]
%\centering
%\includegraphics[scale=0.2]{initial_phi.eps}
%\includegraphics[scale=0.2]{final_phi.eps}
%\\
%\includegraphics[scale=0.2]{initial_cross_section.eps}
%\includegraphics[scale=0.2]{final_cross_section.eps}
%\caption{Reinitialization procedure. Left: Initial level set function (top) and its zero cross
%section (bottom) defining a square domain. Right: Reinitialized level set function (top) and its
%zero cross section (bottom). In particular, the difference in the zero level set between the
%initial and final stages is negligible. In this example, the level difference between adjacent
%cells is not restricted.} \label{fig_ex_reinitialization_square}
%\end{figure}

\begin{figure}[t]
\centering
\includegraphics[scale=0.2]{losange_step.eps}
\includegraphics[scale=0.2]{losange_reinitialized.eps}
\\
\includegraphics[scale=0.2]{losange_initial.eps}
\includegraphics[scale=0.2]{losange_final.eps}
\caption{Reinitialization procedure. Left: Initial level set function (top)
and its zero cross section (bottom) defining a square domain. Right:
Reinitialized level set function (top) and its zero cross section (bottom).
In particular, the difference in the zero level set between the initial and
final stages is negligible. In this example, the level difference between
adjacent cells is not restricted.} \label{fig_ex_reinitialization_square}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=1.0]{reinitialization_error_decrease.eps}
\end{center}
\caption{$L_1$ errors of the reinitialization algorithm in the case of the
adaptive time step (solid line) and the uniform time step (dotted line).}
\label{fig_reinitialization_error_decay}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{reinitialization_initial.eps}
\includegraphics[scale=0.8]{reinitialization_5_iterations.eps} \\
\includegraphics[scale=0.8]{reinitialization_10_iterations.eps}
\includegraphics[scale=0.8]{reinitialization_20_iterations.eps}
\end{center}
\caption{From top-left to bottom-right: Contours of the reinitialized level
set function of example \ref{subsec_Accuracy_Reinitialization} after 0, 5,
10 and 20 iterations. The contours are evenly plotted from -1 to 1 with a
thick line representing the zero contour.} \label{fig_reinitialization}
\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multicolumn{3}{|c|}{} & \multicolumn{5}{|c|}{Finest Resolution}   \\\cline{4-8}
\multicolumn{3}{|c|}{} & $128^2$ & rate & $256^2$ & rate & $512^2$ \\\hline \multirow{4}{0.7
in}{Uniform Refinement} & \multirow{2}{*}{Near Interface}
 & $L_1     $ & $4.36\times10^{-6}$ & 2.92 & $5.77\times10^{-7}$ & 3.02 & $7.12\times10^{-8}$ \\\cline{3-8}
&& $L_\infty$ & $2.16\times10^{-5}$ & 2.74 & $3.24\times10^{-6}$ & 3.26 & $3.38\times10^{-7}$
\\\cline{2-8}
                                       & \multirow{2}{*}{Whole Domain}
 & $L_1     $ & $3.27\times10^{-4}$ & 2.14 & $7.42\times10^{-5}$ & 2.11 & $1.71\times10^{-5}$ \\\cline{3-8}
&& $L_\infty$ & $4.20\times10^{-2}$ & 1.56 & $1.43\times10^{-2}$ & 1.87 & $3.89\times10^{-3}$
\\\hline

\multirow{4}{0.7 in}{Adaptive Refinement} & \multirow{2}{*}{Near Interface}
 & $L_1     $ & $4.36\times10^{-6}$ & 2.94 & $5.70\times10^{-7}$ & 3.00 & $7.14\times10^{-8}$ \\\cline{3-8}
&& $L_\infty$ & $2.16\times10^{-5}$ & 2.87 & $2.96\times10^{-6}$ & 3.09 & $3.48\times10^{-7}$
\\\cline{2-8}
                                         & \multirow{2}{*}{Whole Domain}
 & $L_1     $ & $3.27\times10^{-4}$ & 1.06 & $1.57\times10^{-4}$ & 1.01 & $7.82\times10^{-5}$ \\\cline{3-8}
&& $L_\infty$ & $4.20\times10^{-2}$ & 0.00 & $4.20\times10^{-2}$ & 0.00 & $4.20\times10^{-3}$
\\\hline
\end{tabular}
\end{center}
\caption{Convergence rates for the reinitialization for example
\ref{subsec_Accuracy_Reinitialization}. The initial grid is shown in figure
\ref{fig_reinitialization}. The condition for a node $v_i$ to be
'\textit{near interface}' is chosen as $|\phi(v_i)|<\sqrt{2}\triangle
x_{\text{smallest}}$, where $\triangle x_{\text{smallest}}$ is the size of
the smallest cell. The '\textit{whole domain}' excludes the region near the
kink located at the origin, where accuracy drops to first order.}
\label{tab_reinitialization}
\end{table}

%-------------------------------------------------------------------------------
%
% semi-Lagrangian Method
%
%-------------------------------------------------------------------------------
%\section{Semi-Lagrangian Method\label{sec_semiL}}
%-------------------------------------------------------------------------------
%
% Moving Interface with Given Velocity
%
%-------------------------------------------------------------------------------
\section{Motion Under an Externally Generated Velocity Field}\label{sec_semi_lagrangian}
\subsection{Second Order Accurate Semi-Lagrangian Method\label{Sec_midpoint_rule}}
In the case where the velocity field is externally generated, the level set
equation \eqref{eq_levelset} is linear. In this case, one can use
semi-Lagrangian methods. Semi-Lagrangian schemes are extensions of the
Courant-Isaacson-Rees \cite{Courant:1952:CIR_Method} method for hyperbolic
equations and are unconditionally stable thus avoiding standard CFL
condition of $\Delta t \approx \Delta x_{\rm smallest}$. The general idea
behind semi-Lagrangian methods is to reconstruct the solution by
integrating numerically the equation along characteristic curves, starting
from any grid point $x_i$ and tracing back the departure point $x_d$ in the
upwind direction. Interpolation formulas are then used to recover the value
of the solution at such points. In this work, we use a second order
accurate Semi-Lagrangian method.

Consider the linear advection equation:
\begin{eqnarray}
\phi_t + U\cdot\nabla\phi&=&0, \label{Level_Set_Equation}
\end{eqnarray}
where $U$ is an externally generated velocity field. Then
$\phi^{n+1}(x^{n+1})=\phi^n(x_d)$, where $x^{n+1}$ is any grid node and
$x_d$ is the corresponding departure point from which the characteristic
curve originates. In this work, we use the second order mid-point method
for locating the departure point, as in \cite{Xiu:2001:Semi_Lagrangian_BD}:
$$ \begin{array}{rcl}
\hat{x} &=& x^{n+1} - \frac{\Delta t}2\cdot U^n(x^{n+1}), \\
x_d     &=& x^{n+1} -       \Delta t  \cdot U^{n+\frac12}(\hat{x}),
\end{array}
$$
where we define the velocity at the mid-time step $t^{n+\frac12}$ by a
linear combination of the velocities at the two previous time steps, i.e.
$U^{n+\frac12}=\frac32U^n-\frac12U^{n-1}$. Since $\hat{x}$ and $x_d$ are
not on grid nodes in general, interpolation procedures must be applied to
define $U^{n+\frac12}(\hat{x})$ and $\phi^n(x_d)$. We note that it is
enough to define $U^{n+\frac12}(\hat{x})$ with a multilinear interpolation
\eqref{Bilinear_Interpolation} and $\phi^n(x_d)$ with the quadratic
interpolation described by equations \eqref{Stabilized_Interpolation1} and
\eqref{Stabilized_Interpolation2}: Since a distance function has
discontinuities in its derivative in general, the stabilized quadratic
interpolation is preferred to the Hermite quadratic interpolation.

%-------------------------------------------------------------------------------
% Example : 2D Rotation Test
%-------------------------------------------------------------------------------
\subsection{Test: Rotation in 2D \label{ex_rotation_test_2D}}

Consider a domain $\Omega=[-1, 1]^2$ and a disk of radius $R=.15$ and center initially at $(0,
.75)$, rotating under the divergence free velocity field
$$
\begin{array}{lcr}
u(x,y) &=& -y \\
v(x,y) &=& x \\
\end{array}
$$
The final time $t=2\pi$ is the time when the rotation completes one
revolution. In the simulation, the adaptive refinement is used, and the
time step restriction is $\Delta t= 5\Delta x$. Table
\ref{tab_ex_rotation_test_2D} demonstrates second order accuracy for the
level set as well as for the mass conservation.  We note that we only
consider the grid nodes neighboring the interface in our computation of the
accuracy for the level set function $\phi$ since only those points define
the location of the interface.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Finest &$L^\infty$ error&rate&$L^1 error$&rate&Loss of&rate \\
Resolution& of $\phi$      &    & of $\phi$ &    &volume(\%)&    \\\hline
$32^2$&$7.24\times10^{-2}$&    &$3.11\times10^{-2}$&    &28.51&  \\\hline
$64^2$&$1.78\times10^{-2}$&2.02&$8.86\times10^{-3}$&1.81&7.21 & 1.98   \\\hline
$128^2$&$4.52\times10^{-3}$&1.97&$2.13\times10^{-4}$&2.05&1.78& 2.01   \\\hline
$256^2$&$1.13\times10^{-3}$&1.99&$5.56\times10^{-4}$&1.93&0.45& 1.98   \\\hline
$512^2$&$2.85\times10^{-4}$&2.00&$1.38\times10^{-4}$&2.01&0.11& 2.03   \\\hline
$1024^2$&$7.14\times10^{-5}$&2.00&$3.46\times10^{-5}$&2.00&0.03& 1.87   \\\hline
$2048^2$&$1.78\times10^{-5}$&2.00&$8.64\times10^{-6}$&2.00&0.007& 2.01   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rates for example \ref{ex_rotation_test_2D}
\label{tab_ex_rotation_test_2D}}
\end{table}

%-------------------------------------------------------------------------------
% Example : Vortex Test Test
%-------------------------------------------------------------------------------
\subsection{Test: Vortex in 2D \label{ex_large_deformation_test_2D}}
In this example, we test our level set implementation on the more
challenging flow proposed by Bell \etal \cite{Bell_Colella_Glaz}: Consider
a domain $\Omega=[0, 1]^2$ and a disk of radius $.15$ and center $(.5,
.75)$ as the initial zero level set contour. The level set is then deformed
under the divergence free velocity field $U=(u,v)$ given by:
$$
\begin{array}{rcr}
u(x,y) &=& -\sin^2(\pi x)\sin(2\pi y) \\
v(x,y) &=&  \sin^2(\pi y)\sin(2\pi x) \\
\end{array}
$$
The disk is deformed forward until $t=1$ and then backward to the original
shape using the reverse velocity field with a time step restriction of
$\triangle t=5\cdot\triangle x_{\text{smallest}}$.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}   \\\cline{2-8}
                  & $64^2$ & rate & $128^2$ & rate & $256^2$ & rate & $512^2$ \\\hline
$L_1     $ error of $\phi$&$9.58\TT{-3}$&2.80&$1.38\TT{-3}$&2.02&$3.41\TT{-4}$&2.08&$8.09\TT{-5}$
\\\hline $L_\infty$ error of
$\phi$&$1.83\TT{-2}$&1.57&$6.17\TT{-3}$&1.08&$2.91\TT{-3}$&1.39&$1.11\TT{-3}$ \\\hline Volume Loss
&$4.48       $&2.36&$0.874      $&1.40&$0.331      $&1.79&$0.0954     $ \\\hline Max number of
nodes   &$1045       $&1.10&$2243       $&1.10&$4815       $&1.09&$10256      $ \\\hline Min number
of nodes   &$439        $&1.07&$924        $&0.99&$1831       $&1.01&$3679       $ \\\hline
time(sec) & 1.420 & 2.29 & 6.96 & 2.17 &31.4 & 2.13 & 138 \\
\hline minimum memory(MB) & 0.0448 & 1.07 & 0.0943 & .97 & 0.185 & 1.00 & 0.371 \\
\hline maximum memory(MB) & 0.101 & 1.30 & 0.248 & 1.06 & 0.517 & 1.03 & 1.06 \\ \hline
\end{tabular}
\end{center}
\caption{Convergence rates for example \ref{ex_large_deformation_test_2D}. The memory
requirement increases linearly with effective resolution since most of the computational resources
is focused near the one-dimensional interface, i.e. our method is an efficient implementation of
local level set methods. The computational time increases quadratically with effective resolution,
since the number of nodes is doubled and the time step is halved.} \label{tab_vortex}
\end{table}

Table \ref{tab_vortex} demonstrate second order accuracy for $L_1$ error of
$\phi$ and volume of loss, and linear increase in the maximum/minimum
number of nodes. Note that the uniform grid of resolution $512^2$ requires
about 25 times more nodes than the adaptive grid with the same resolution.
Although second order accuracy was achieved in both the maximum and average
norms in the previous example, the convergence rate of the maximum error is
oscillating between one and two. This is due to the fact that as the
interface deforms, some part of the interface are under resolved. Figure
\ref{fig_vortex_test} illustrates this at $t=1$: Here, the tail of the
interface is not resolved accurately. This deterioration in accuracy was
also reported in \cite{Olsson_Kreiss:2005}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{vortex_flow_at_one.eps}
\includegraphics[scale=0.7]{vortex_flow_at_two.eps}
%%includegraphics[scale=0.7]{volume_loss.eps}
\end{center}
\caption{Contours of the zero level sets for example
\ref{ex_large_deformation_test_2D} with effective resolutions of $64^2$,
$128^2$, $256^2$ and $512^2$ at $t=1$ (left) and $t=2$ (right). The colors
red, green, blue represent the difference in the interface location between
the resolutions $64^2$ and $128^2$, $128^2$ and $256^2$, $256^2$ and
$512^2$, respectively.} \label{fig_vortex_test}
\end{figure}

Figure \ref{large_deformation_test_2D} illustrates the evolution of the
interface location initially (left), at $t=6$ (center) and when the
interface is fully rewinded (right). This example illustrates the ability
of the present method to accurately capture the evolution of an interface
undergoing large deformations and the ability to preserve mass effectively
(mass loss $\approx.3\%$).

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{Large_Deformation_2D_T6_Initial.eps}
\includegraphics[scale=0.25]{Large_Deformation_2D_T6_Mid.eps}
\includegraphics[scale=0.25]{Large_Deformation_2D_T6_Final.eps}
\end{center}
\caption{Level set evolution at $t=0$ (left), $t=3$ (center) and $t=6$ (right). The effective
resolution is $2048^2$ and the mass is conserved within .3\%} \label{large_deformation_test_2D}
\end{figure}

%-------------------------------------------------------------------------------
% Example : 2D Zalesak's Disk
%-------------------------------------------------------------------------------
%\subsection{Zalesak's Disk in 2D \label{ex_zalesak_2d}}
%In this example, we test our method on the Zalesak's disk problem in 2D. Consider a domain $\Omega=[0, 100]^2$ and a disk of radius $15$ and center
%$(50, 75)$ that is slotted of width $5$ and height $25$ as the initial zero level set contour. The slotted disk is rotated around $(50,50)$ with velocity field $\frac\pi{314}(50-y,x-50)$. After one revolution that happens when $T=628$, the loss of volume was measured in Table \ref{tab_ex_zalesak_2d}. The volume losses of 0.8\% in $128^2$ and 0.26\% in $256^2$ are similiar to the results of Particle Level Set method in \cite{En} of 0.31\% in $100^2$ and 0.20\% in $200^2$.

### 2012-11-19 doc/levelset/level_set_method_on_quadtree_JCP_rebuttal.tex

%\begin{table}
%\begin{center}
%\begin{tabular}{|c|c|c|}
%\hline
%The most    & Loss of & rate \\
%refined grid& Volume  &      \\\hline
%$ 64^2$&4.60  \% & 2.51  \\\hline
%$128^2$&0.808 \% & 2.51  \\\hline
%$256^2$&0.259 \% & 1.64  \\\hline
%$512^2$&0.0468\% & 2.47  \\\hline
%\end{tabular}
%\end{center}
%\caption{Convergence rate for example \ref{ex_zalesak_2d} \label{tab_ex_zalesak_2d}}
%\end{table}

%-------------------------------------------------------------------------------
% Example : Reinitialization_3D
%-------------------------------------------------------------------------------
%\subsection{Reinitialization in 3D \label{ex_reinitialization_3d}}
%In this section, we test the convergence rate of the reinitialization algorithm in octree. In $[0,1]^3$, a signed distance function to the sphere of center $(0.35,0.35,0.35)$ and of radius $0.15$ is reinitialized in an octree depicted in Figure \ref{fig_ex_reinitialization_3d}. In the next refinement, the automatic grid generation algorithm is used. Table \ref{tab_ex_reinitialization_3d} shows that the reinitialization method is 3rd order accurate. As shown in figure \ref{fig_ex_reinitialization_3d}, the reinitialization in a coarse grid of $16^3$ could not be noticed visually.

%\begin{figure}[t]
%\centering
%\includegraphics[scale=0.4]{octree_before_reinitialization.eps}
%\includegraphics[scale=0.4]{octree_after_reinitialization.eps}
%\caption{Level sets on a octree before and after the reinitialization in $16^3$ grid for Example \ref{ex_reinitialization_3d}}
%\label{fig_ex_reinitialization_3d}
%\end{figure}

%\begin{table}
%\begin{center}\begin{tabular}{|c|c|c|c|c|c|c|c|}
%\hline
%Effective  &&&&&&&\\
%Resolution & Error in $L_\infty$ & Order & Error in $L_1$ & order & Volume Loss & \#nodes &\#cells %\tabularnewline \hline
%$  32^3$ & $3.63\times 10^{-4}$ &      & $1.26\times 10^{-4}$ &      & - 2.1\% & 3541 & 3137 \tabularnewline \hline
%$  64^3$ & $9.89\times 10^{-5}$ & 1.87 & $1.54\times 10^{-5}$ & 3.03 & -0.54\% & 12059 & 11257 \tabularnewline \hline
%$ 128^3$ & $3.18\times 10^{-5}$ & 1.63 & $2.22\times 10^{-6}$ & 2.79 & -0.13\% & 44955 & 42777 \tabularnewline \hline
%$ 256^3$ & $3.46\times 10^{-6}$ & 3.19 & $2.59\times 10^{-7}$ & 3.09 & -0.034\% & 176236 & 168881 \tabularnewline \hline
%$ 512^3$ & $5.05\times 10^{-7}$ & 2.77 & $3.36\times 10^{-8}$ & 2.94 & -0.0084\% & 701711 & 674089 \tabularnewline \hline
%\end{tabular}
%\end{center}
%\caption{Convergence rate of the reinitialization algorithm in 3D for  example \ref{ex_reinitialization_3d}
%\label{tab_ex_reinitialization_3d}}
%\end{table}

%-------------------------------------------------------------------------------
% Example : 3D Rotation Test
%-------------------------------------------------------------------------------
\subsection{Test: Rotation in 3D \label{ex_rotation_test_3D}}

Consider a domain $\Omega=[-2, 2]^3$ and a sphere of radius $R=.5$ and
center initially at $(0,1,0)$, rotating under the divergence free velocity
field
$$
\begin{array}{lcr}
u(x,y,z) &=& -y \\
v(x,y,z) &=& x  \\
w(x,y,z) &=& 0
\end{array}
$$
The simulation is run until $t=2\pi$, when the rotation completes one
revolution. In the simulation, the adaptive refinement is used, and the
time step restriction is $\Delta t=6\Delta x_{\rm smallest}$. Table
\ref{tab_ex_rotation_test_3D} demonstrates second order accuracy for the
level set as well as for the mass conservation.  We note that we only
consider the grid nodes neighboring the interface in our computation of the
accuracy for the level set function $\phi$. Figure
\ref{fig_rotating_sphere} shows the adaptive grid for the rotating sphere.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}\\\cline{2-8}
                  & $32^3$ & rate & $64^3$ & rate & $128^3$ & rate & $256^3$ \\\hline
$L_1$      error of $\phi$ & $6.86\TT{-2}$ & 1.88 & $1.87\TT{-2}$ & 1.99 & $4.70\TT{-3}$ & 1.99 &
$1.18\TT{-3}$ \\\hline $L_\infty$ error of $\phi$ & $1.76\TT{-1}$ & 2.02 & $4.35\TT{-2}$ & 2.02 &
$1.07\TT{-2}$ & 2.02 & $2.65\TT{-3}$ \\\hline Volume loss(\%)            & 23.1 & 2.16 & 5.14 &
2.12 & 1.18 & 2.07 & 0.282 \\\hline
\end{tabular}
\end{center}
\caption{Convergence rates for the interface's location for example \ref{ex_rotation_test_3D}.}
\label{tab_ex_rotation_test_3D}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{rotating_sphere_initial.eps}
\includegraphics[scale=0.8]{rotating_sphere_a_quarter.eps}\\
\includegraphics[scale=0.8]{rotating_sphere_a_half.eps}
\includegraphics[scale=0.8]{rotating_sphere_final.eps}
\end{center}
\caption{Evolution of the interface for example \ref{ex_rotation_test_3D}: Initial data (top-left),
interface after a quarter turn (top-right), interface after a half turn (bottom-left) and final
location (bottom-right). The finest resolution is $128^3$.} \label{fig_rotating_sphere}
\end{figure}

%-------------------------------------------------------------------------------
% Example : Enright's Test in 3D
%-------------------------------------------------------------------------------
\subsection{Enright's Test in 3D \label{ex_enright_test}}
We consider the test proposed in Enright \etal \cite{Enright:2002:PLS}: A
sphere of center $(0.35,0.35,0.35)$ and radius $0.15$ in the domain of
$[0,1]^3$ is deformed under the following divergence free velocity field:
$$
\begin{array}{lcr}
u(x,y,z) &=& 2\sin^2(\pi x)\sin(2\pi y)\sin(2\pi z) \\
v(x,y,z) &=& -\sin^2(\pi y)\sin(2\pi x)\sin(2\pi z) \\
w(x,y,z) &=& -\sin^2(\pi z)\sin(2\pi x)\sin(2\pi y)
\end{array}
$$
forward in time and then backward to its original shape with the reversed velocity. Figure
\ref{fig_enrights_test_until_one} illustrates the interface motion with a time step restriction of
$\triangle t=5\cdot\triangle x_{\text{smallest}}$. We note that, in the simulation with an
effective resolution of $512^3$, minimum the number of nodes used was $685220$ and the maximum was
$2606710$. In contrast, the number of nodes in the case of a uniform grid with the same resolution,
the number of nodes would be about 50 times larger. The volume loss is 3.21\% for an effective
resolution of $256^3$ and 0.739\% for an effective resolution of $512^3$. Figure
\ref{fig_enrights_test_refinement} compares the interface evolution with an effective resolution of $128^3$, $256^3$ and $512^3$. Table \ref{Enright_Table1} describes the memory and CPU requirements, and table \ref{Enright_Table2} describes the volume loss and the accuracy of the interface location after reconvering the original shape.

The Enright's test is a canonical example to test the amount of numerical dissipation of level set
methods. The particle level set method reported 2.6\% volume loss on a $100^3$ uniform grid
together with Lagrangian particles \cite{Enright:2002:PLS}. Our results show that we obtain a loss
of mass of $.74\%$ in the case of a $512^3$ effective resolution, which corresponds to a $137^3$
uniform grid in term of number of nodes. The particle level set was further improved in
\cite{Enright:2002:AAR} using octree data structures in addition to particles. Although
\cite{Enright:2002:AAR} does not report any quantitative results, we find that our result for the
Enright's test is visually comparable to that obtained in \cite{Enright:2002:AAR} for the same
effective resolution and compares favorably with the results in \cite{Hieber:Particle_Level_Set}.

\begin{table}
\begin{center}\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
  \hline
  Finest     &\multirow{2}{*}{Time}
             &\multirow{2}{*}{Rate}
             & Min \#
             &\multirow{2}{*}{Rate}
             & Max \#
             &\multirow{2}{*}{Rate}
             & Min
             &\multirow{2}{*}{Rate}
             & Max
             &\multirow{2}{*}{Rate}  \\
  Resolution & & & nodes & & nodes & & Memory & & Memory &  \\  \hline
  $128^3$ & 237.5 &  & 44943 &  & 133308 &  & 4.54 &  & 13.7 &  \\  \hline
  $256^3$ & 2214 & 3.23 & 173637 & 1.95 & 598264 & 2.17 & 17.6 & 1.96 & 61.5 & 2.17 \\  \hline
  $512^3$ & 19521 & 3.14 & 685220 & 1.98 & 2606710 & 2.12 & 69.6 & 1.98 & 268 & 2.12 \\  \hline

  \end{tabular}\caption{Memory and CPU requirements for example \ref{ex_enright_test}. The memory
requirement increases quadratically with effective resolution since most of the computational
resources is focused near the two-dimensional interface, i.e. our method is an efficient
implementation of local level set methods. The computational time increases cubically with
effective resolution, since the number of nodes is multiplied by four and the time step is halved.}
  \label{Enright_Table1}
  \end{center}  \end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
  Finest Resolution & Volume Loss & Rate & $L^1     $ error of $\phi$ & Rate
                                         & $L^\infty$ error of $\phi$ & Rate \\\hline
  $128^3$ & 16.02\% &  & $1.96 \times 10^{-2}$ &  & $1.54 \times 10^{-1}$&\\\hline
  $256^3$ & 3.21\% & 2.32 & $2.83 \times 10^{-3}$ & 2.79 & $1.06 \times 10^{-1}$& .88\\\hline
  $512^3$ & .739 \% & 2.12 & $4.38 \times 10^{-4}$ & 2.69 & $5.74 \times 10^{-3}$& 7.52\\\hline
\end{tabular}\caption{Convergence rates for the interface's location for example \ref{ex_enright_test}.}
\label{Enright_Table2}
\end{center}\end{table}

We note that the jump in rate in table \ref{Enright_Table2} can be explained by the lack of
resolution for describing the developing thin film. This is related to the Nyquist-Shannon sampling
theorem that states that in order to fully reconstruct a signal the sampling frequency should be at
least twice the signal bandwidth. In our case, the fact that the thin film is under-resolved
prevents subcell resolution of the reinitialization scheme. For higher resolutions, we would expect
second-order accuracy.

\begin{figure}
\begin{center}
\includegraphics[scale=0.3]{enrights_test_512_until_1.eps}
\end{center}
\caption{Evolution of the interface for the Enright's test with finest resolution of $512^3$.}
\label{fig_enrights_test_until_one}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.3]{image_0001_resolution_128.eps}\includegraphics[scale=0.3]{image_0001_resolution_256.eps}\includegraphics[scale=0.3]{image_0001_resolution_512.eps}\\
\includegraphics[scale=0.3]{image_0002_resolution_128.eps}\includegraphics[scale=0.3]{image_0002_resolution_256.eps}\includegraphics[scale=0.3]{image_0002_resolution_512.eps}\\
\end{center}
\caption{Effect of refinement on the Enright's test: The top figures correspond to the interface
fully stretched and the bottom figures correspond to the interface rewinded to the original sphere.
The finest resolutions are $128^3$ (left), $256^3$ (center) and $512^3$ (right).}
\label{fig_enrights_test_refinement}
\end{figure}

%-------------------------------------------------------------------------------
%
% Moving Interface with Curvature Dependent Speed
%
%-------------------------------------------------------------------------------
\section{Motion in the Normal Direction and Curvature Driven Flow}
The equation describing an interface propagating in its normal direction and under its mean
curvature is given by \cite{osher:1988:originallevelset}:
\begin{equation}
\phi_t+(\alpha-\beta\kappa)|\nabla\phi|=0, \label{Motion_Normal_Curvature}
\end{equation}
where $\kappa$ is the mean curvature of the interface $\kappa=\nabla\cdot\left(
\nabla\phi/|\nabla\phi| \right)$. The coefficients $\alpha$ and $\beta\ge0$ control the magnitude
of the speed in the normal direction and the strength of the curvature dependence, respectively.
The case where $\beta<0$ is ill-posed and therefore we do not consider it here.

\subsection{Motion in the Normal Direction}
First, we discuss the case when $\beta=0$. Using the second order one-sided derivatives described
in section \ref{sec_finite_differences} and discretizing the Hamiltonian using a Godunov scheme, we
semi-discretize the equation as :
$$\frac{d\phi}{dt} + \alpha\cdot H_G(\phi)=0$$
where the Godunov Hamiltonian $H_G$ is defined as:
$$H_G(\phi) =
\begin{cases}
\sqrt{\max(|(D_x^+\phi)^-|^2,|(D_x^-\phi)^+|^2)+\max(|(D_y^+\phi)^-|^2,|(D_y^-\phi)^+|^2)}&\text{if $\alpha>0$}\\
\sqrt{\max(|(D_x^+\phi)^+|^2,|(D_x^-\phi)^-|^2)+\max(|(D_y^+\phi)^+|^2,|(D_y^-\phi)^-|^2)}&\text{otherwise}
\end{cases}$$
This equation is discretized in time using the second order TVD Runge-Kutta method (see
\cite{Shu:1988:ENO,liu:1996:WENO}):
\begin{align}
\frac{\tilde{\phi}^{n+1}-       \phi ^n}{\triangle t} &+ \alpha\cdot H_G\left(\phi^n\right)=0 \\
\frac{\tilde{\phi}^{n+2}-\tilde{\phi}^n}{\triangle t} &+ \alpha\cdot H_G\left(\tilde{\phi}^{n+1}\right)=0 \\
\phi^{n+1}&=\frac{\phi^n+\tilde{\phi}^{n+2}}2
\end{align}

Table \ref{tab_shrinking_circle} illustrates that the method described above is second order
accurate in both the maximum and the average norms for smooth data. In the case where the interface
presents sharp corners, figure \ref{fig_shrinking_and_expading_square} illustrates that the method
converges to the correct viscosity solution \cite{osher:1988:originallevelset}.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}\\\cline{2-8}
                  & $64^2$ & rate & $128^2$ & rate & $256^2$ &rate & $512^2$ \\\hline
$L_1$      error of
$\phi$&$1.46\TT{-3}$&2.02&$3.58\TT{-4}$&2.00&$8.56\TT{-4}$&1.98&$2.26\TT{-5}$\\\hline $L_\infty$
error of $\phi$&$2.77\TT{-3}$&2.04&$6.72\TT{-4}$&1.95&$1.73\TT{-4}$&1.99&$4.36\TT{-5}$\\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for a circle shrinking with unit normal velocity. consider a domain of
$[-2,2]^2$ and an interface initially described by a circle centered at the origin with radius
$R=1$. The interface is evolved until $t=0.5$.} \label{tab_shrinking_circle}
\end{table}
\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{initial_square_new.eps}
\includegraphics[scale=0.5]{shrinking_square_point_5.eps}
\includegraphics[scale=0.5]{shrinking_square_1.eps}\\
\includegraphics[scale=0.5]{initial_square_new.eps}
\includegraphics[scale=0.5]{expanding_square_point_5.eps}
\includegraphics[scale=0.5]{expanding_square_1.eps}
\end{center}
\caption{Shrinking square in the first row, and expanding square in the second row}
\label{fig_shrinking_and_expading_square}
\end{figure}

\subsection{Adding Motion by Mean Curvature}
Now we discuss the case when $\beta>0$. The curvature term can be
discretized explicitly or implicitly. In the case where the curvature term
is discretized explicitly, the corresponding time step restriction of
$\Delta t \approx \Delta x^2$ is too stringent to be practical since it
would be constrained by the size of the smallest grid cell in the grid. In
\cite{Smereka:2003:Curvature}, Smereka proposed an implicit discretization
of the curvature term in the case of uniform grids: Using the following
operator splitting:
$$\kappa|\nabla\phi|=\Delta \phi -
\frac{\nabla\phi}{|\nabla\phi|}\nabla(|\nabla\phi|),$$ equation \eqref{Motion_Normal_Curvature} is
discretized as:
$$\frac{\phi^{n+1}-\phi^n}{\triangle t} +\alpha H_G(\phi^n) = \beta\Delta \phi^{n+1}
- \beta\frac{\nabla\phi^n}{|\nabla\phi^n|}\cdot\nabla(|\nabla\phi^n|).$$

In this work, we used a Backward Euler step to treat the linear term, and a
Forward Euler step for the nonlinear term. The derivatives $\Delta$ and
$\nabla$ are discretized by the central finite differences described in
section \ref{sec_finite_differences}. Discretizing implicitly the Laplacian
requires a linear system that we solve using the supra convergent method
presented in Min, Gibou and Ceniceros \cite{Min::Poisson_AMR_Second_Order}.
As noted in \cite{Smereka:2003:Curvature}, the semi-implicit discretization
on the curvature term allows for a big time step, so that the time step
restriction is that of the convection part, i.e.
\begin{equation}
\Delta t=\frac{\Delta x_{\text{smallest}}}{\alpha\cdot\text{\# dimensions}},
\end{equation} where $\text{\# dimensions}$ is the number of dimensions.
